---
title: "Anti-Trans Laws and Mental Health: Tables and Figures"
format: html
editor: visual
execute: 
  echo: false
  warning: false
  message: false
---

# setup

```{r}
#| label: loading in data and packages

library(tidyverse)
library(here)
library(did)
library(patchwork)
load(here("data_clean", "datasets.Rdata"))
load(here("did_results", "coeffs.Rdata"))
load(here("data_clean", "states_sf.rda"))
load(here("data_clean", "all_laws.Rdata"))

```

```{r}
#| label: data wrangling

# making some relevant variables
df_queer <- map(df_queer, ~drop_na(.x,
                                   anxious, worry, interest, down) |> 
                  mutate(phq4 = anxious + worry + interest + down - 4, #HPS codes these from 1-4; we need them added as though they are 0-3
                          group = case_when(trans_gnc == 1 ~ treatment_period,
                                              trans_gnc == 0 ~ 0), #`did` package convention: comparison group's group is 0
                         coldeg = case_match(educ, 
                                             c(1:4) ~ 0,
                                             c(5:7) ~ 1,
                                             .default = NA_integer_),
                         nhw = case_match(as.numeric(race_rc),
                                          1 ~ 1,
                                          c(2:4) ~ 0,
                                          .default = NA_integer_),
                         age = year(startdate) - birth_year))

#using the same process on the trans datasets, for later
df_trans <- map(df_trans, ~drop_na(.x,
                                   anxious, worry, interest, down) |> 
                  mutate(phq4 = anxious + worry + interest + down - 4,
                         group = case_when(state_type == "treated" ~ treatment_period,
                                           state_type == "notyettreated" ~ 0),
                         coldeg = case_match(educ, 
                                             c(1:4) ~ 0,
                                             c(5:7) ~ 1,
                                             .default = NA_integer_),
                         nhw = case_match(as.numeric(race_rc),
                                          1 ~ 1,
                                          c(2:4) ~ 0,
                                          .default = NA_integer_),
                         age = year(startdate) - birth_year))

```

# Main Results

## Table 1. Sample Characteristics

```{r}
#| label: sample characteristics
# the queer comparison is the main result; making trans vs. trans a supportive supplemental argument

sc <- map(df_queer, ~group_by(.x, trans_gnc, post_period) |> 
  summarize(n = n(),
            mean_phq4 = weighted.mean(phq4, w = pweight),
            sd_phq4 = sd(phq4),
            mean_age = weighted.mean(age, w = pweight),
            sd_age = sd(age),
            prop_nhw = weighted.mean(nhw, w = pweight),
            prop_coldeg = weighted.mean(coldeg, w = pweight),
            prop_fatbirth = weighted.mean(genid_birth - 1, w = pweight)) |> 
  ungroup())
  
sc <- sc |> 
  tibble() |> 
  mutate(analysis = names(sc)) |> 
  unnest(sc) |> 
  pivot_longer(cols = c(n:prop_fatbirth),
               names_to = "stat",
               values_to = "value") |> 
  mutate(trans_gnc = case_match(trans_gnc,
                                0 ~ "queer",
                                1 ~ "trans"),
         post_period = case_match(post_period,
                                  0 ~ "pre",
                                  1 ~ "post")) |> 
  pivot_wider(id_cols = stat,
              names_from = c(trans_gnc, post_period, analysis),
              values_from = value,
              names_sep = "_")

write_csv(sc, file = here("tables_figs", "desctable.csv"))
```

## Figure 1. States Visualization

```{r}
analysis_states <- df_queer |> map(~group_by(.x, state) |> 
  summarize(n = n()) |> 
  pull(state))

mapdata <- states_sf |> 
  mutate(first = st %in% analysis_states$first,
         bathroom = st %in% analysis_states$bathroom,
         medical = st %in% analysis_states$medical,
         sports = st %in% analysis_states$sports)

 hcl.colors(4, palette = "Pastel 1")
 # "#FFC5D0" "#D4D8A7" "#99E2D8" "#D5D0FC"

map_b <- ggplot(data = mapdata) +
  geom_sf(aes(fill = bathroom)) +
  scale_fill_manual(values = c("grey95", "#99E2D8")) +
  theme_void() +
  labs(title = "First Bathroom Law") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

map_m <- ggplot(data = mapdata) +
  geom_sf(aes(fill = medical)) +
  scale_fill_manual(values = c("grey95", "#FFC5D0")) +
  theme_void() +
  labs(title = "First Medical Law") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

map_s <- ggplot(data = mapdata) +
  geom_sf(aes(fill = sports)) +
  scale_fill_manual(values = c("grey95", "#D4D8A7")) +
  theme_void() +
  labs(title = "First Sports Law") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

map_f <- ggplot(data = mapdata) +
  geom_sf(aes(fill = first)) +
  scale_fill_manual(values = c("grey95", "#D5D0FC")) +
  theme_void() +
  labs(title = "First Law Overall") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))


map_b + map_m + map_s + map_f + plot_layout(nrow = 2)


```

## Figure 2. PHQ-4 Trends Visualization

```{r}

```

## Figure 3. DID Results

# Supplemental

## eTable 1. Post-Treatment ATTs vs Pre-Treatment ATTs

```{r}
#| label: post vs. pre ratios
coeffs$sports |> 
  filter(t < -1) |> 
  group_by(modelspec) |> 
  summarize(overall_att = mean(overall_att),
            avg_pre = mean(abs(att))) |> 
  mutate(ratio = overall_att / avg_pre)

coeffs$first |> 
  filter(t < -1) |> 
  group_by(modelspec) |> 
  summarize(overall_att = mean(overall_att),
            avg_pre = mean(abs(att))) |> 
  mutate(ratio = overall_att / avg_pre)

```

## eTable 2. Sample Construction Process

constructed in Word

## eTable 3. Sample Comparisons

## eTable 4. Alternate Specifications

## eTable 5. Sample Characteristics: Trans Sample

## eFigure 1. Panel Structure of Data
