---
title: "Anti-Trans Laws and Mental Health: Tables and Figures"
format: pdf
editor: visual
execute: 
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup - loading in data and packages

library(tidyverse)
library(here)
library(did)
library(patchwork)
library(sf)

load(here("data_clean", "datasets.Rdata"))
load(here("did_results", "coeffs.Rdata"))
load(here("data_clean", "states_sf.rda"))
load(here("data_clean", "all_laws.Rdata"))
load(here("data_clean", "hps_data.Rdata"))

```

# Main Results

## Table 1. Sample Characteristics

```{r}
#| label: sample characteristics
# main results are the first laws, queer comparison group, fully cont.


scq <- df_queer$first |> 
  group_by(trans_gnc) |> 
  summarize(n = n(),
            mean_phq4 = weighted.mean(phq4, w = pweight),
            sd_phq4 = sd(phq4),
            mean_age = weighted.mean(age, w = pweight),
            sd_age = sd(age),
            prop_nhw = weighted.mean(nhw, w = pweight),
            prop_coldeg = weighted.mean(coldeg, w = pweight),
            prop_fatbirth = weighted.mean(genid_birth - 1, w = pweight)) |> 
  mutate(group = case_match(trans_gnc,
                            1 ~ "Treatment Group",
                            0 ~ "Queer Reference Group")) |> 
  select(group, n:prop_fatbirth)

sct <- df_trans$first |> 
  group_by(state_type) |> 
  summarize(n = n(),
            mean_phq4 = weighted.mean(phq4, w = pweight),
            sd_phq4 = sd(phq4),
            mean_age = weighted.mean(age, w = pweight),
            sd_age = sd(age),
            prop_nhw = weighted.mean(nhw, w = pweight),
            prop_coldeg = weighted.mean(coldeg, w = pweight),
            prop_fatbirth = weighted.mean(genid_birth - 1, w = pweight)) |> 
  filter(state_type == "notyettreated") |> 
  mutate(group = case_match(state_type,
                            "notyettreated" ~ "Trans Reference Group")) |> 
  select(group, n:prop_fatbirth)

sc <- bind_rows(scq, sct) |> 
  mutate(phq4 = paste0(round(mean_phq4, digits = 2), " (", round(sd_phq4, digits = 2), ")"),
         age = paste0(round(mean_age, digits = 2), " (", round(sd_age, digits = 2), ")")) |>
  select(group, n, phq4, age, starts_with("prop")) |>
  mutate(across(starts_with("prop"), ~round(.x, digits = 3)),
         n = as.integer(n)) |> 
  rename("PHQ-4 Mean (SD)" = phq4,
         "Age Mean (SD)" = age,
         "Proportion non-Hispanic White" = prop_nhw,
         "Proportion with a College Degree" = prop_coldeg,
         "Proportion Assigned Female at Birth" = prop_fatbirth) |> 
  mutate(across(everything(), as.character)) |> 
  pivot_longer(cols = !group,
               names_to = "stat",
               values_to = "value") |> 
  pivot_wider(id_cols = stat,
              names_from = group,
              values_from = value) |> 
  rename("Statistic" = stat) |> 
  select("Statistic", "Treatment Group",
         "Queer Reference Group", "Trans Reference Group")

write_csv(sc, file = here("tables_figs", "desctable.csv"))


knitr::kable(sc)
```

## Figure 1. States Visualization

```{r}
#| label: map visualization
analysis_states <- all_laws |> 
  select(state, date) |> 
  group_by(state) |> 
  slice_min(date, with_ties = F) |> 
  arrange(date) |> 
  ungroup() |> 
  right_join(tibble(state = state.abb)) |> 
  mutate(period = case_when(is.na(date) ~ "never",
                          date < mdy("July 21, 2021") ~ "before",
                          date >= mdy("July 21, 2021") &
                            date <= mdy("October 30, 2023") ~ "during",
                          date > mdy("October 30, 2023") ~ "after"
                            )) |> 
  select(state, period)

mapdata <- states_sf |> 
  left_join(analysis_states, by = c("st" = "state")) |> drop_na(period)

map <- ggplot(data = mapdata) +
  geom_sf(aes(fill = period)) +
  scale_fill_manual(values = hcl.colors(4, palette = "Pastel 1"),
                    labels = c("After Study Period", "Before Study Period", "During Study Period", "Never (as of 4-5-2026)")) +
  theme_void() +
  labs(fill = "First anti-trans law passed...") +
  theme(legend.position = "bottom", legend.direction = "vertical",
        legend.title.position = "left", legend.key.size = unit(1.2, "char"))


ggsave(plot = map,
       filename = here("tables_figs", "lawsmap.png"),
       dpi = 300,
       height = 7,
       width = 8,
       units = "in")

map
```

## Figure 2. PHQ-4 Trends Visualization

```{r}
#| label: phq-4 visualization

# getting a list of states included in the analysis
include_states <- analysis_states |> 
  filter(period == "during") |> 
  pull(state)

# making the picture
phq4_plot <- hps_data |> 
  filter(state %in% include_states) |> 
  drop_na(queer, trans_gnc) |> 
  mutate(phq4 = anxious + worry + interest + down - 4,
         group = fct(case_when(queer == 0 & trans_gnc == 0 ~ "Cisheterosexual",
                           queer == 1 & trans_gnc == 0 ~ "Cis LGBQ+",
                           queer == 1 & trans_gnc == 1 ~ "Trans/GNC"), levels = c("Trans/GNC", "Cis LGBQ+",                                                           "Cisheterosexual"))) |> 
  group_by(startdate, group) |> 
  summarize(mean_phq = weighted.mean(phq4, w = pweight, na.rm = T)) |> 
  ggplot(aes(x = startdate, y = mean_phq)) +
  geom_line(aes(group = group, color = group)) +
  scale_color_manual(values = hcl.colors(3, palette = "SunsetDark")) +
  scale_x_date(date_labels = "%b %Y") +
  theme_minimal() +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  labs(x = "Date",
       y = "Survey-Weighted Average PHQ-4 Score",
       color = "Group")
  
# saving
ggsave(filename = 
         here("tables_figs", "phq4plot.png"),
       plot = phq4_plot,
       dpi = 300,
       width = 6,
       height = 5,
       units = "in")

phq4_plot
```

## Figure 3. DID Results, Overall

```{r}
mainresults <- coeffs$first |> 
  filter(str_detect(modelspec, "nsg_cov")) |> 
  mutate(comparison_g = case_when(
           str_detect(modelspec, "queer") ~ "Cis LGBQ+ People in\nTreated States",
           str_detect(modelspec, "trans") ~ "Trans People in\nNot-Yet-Treated States"))

overall_did_plot <- mainresults |> 
  ggplot(aes(y = comparison_g)) +
  geom_point(aes(x = overall_att)) +
  geom_segment(aes(x = overall_att + 1.96*overall_se, xend = overall_att - 1.96*overall_se)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(x = "Overall ATT (points on the PHQ-4)",
       y = "Difference-in-Differences\nComparison Group")

ggsave(filename = here("tables_figs",
                       "overall_did_plot.png"),
       plot = overall_did_plot,
       dpi = 300,
       width = 6,
       height = 3,
       units = "in")

overall_did_plot
```

## Figure 4: Dynamic ATTs

```{r}
dynamic_did_plot <- mainresults |> 
  mutate(pre_period = t < -1) |> 
  ggplot(aes(x = t, color = comparison_g,
             alpha = pre_period)) +
  geom_point(aes(y = att)) +
  geom_segment(aes(y = att + 1.96*se, yend = att - 1.96*se)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = -1) +
  facet_wrap("comparison_g") +
  scale_color_manual(values = c("#ABB150", "#ACA2EC")) +
  theme_minimal() +
  scale_alpha_manual(values = c(1, .5)) +
  labs(x = "Waves Since Treatment",
       y = "Dynamic ATTs (points on the PHQ-4)",
       title = "Comparison Group") +
  theme(legend.position = "none",
        panel.background = element_rect(color = "black"),
        plot.title = element_text(hjust = .5,
                                  size = 12))

ggsave(filename = here("tables_figs",
                       "dynamic_did_plot.png"),
       plot = dynamic_did_plot,
       dpi = 300,
       width = 7,
       height = 4.5,
       units = "in")

dynamic_did_plot
```

# Supplemental Material

## eTable 1. Post-Treatment ATTs vs Pre-Treatment ATTs

```{r}
#| label: post vs. pre ratios
ratios <- coeffs$first |> 
  filter(str_detect(modelspec, "nsg_cov")) |> 
  filter(t < -1) |> 
  group_by(modelspec) |> 
  summarize(overall_att = mean(overall_att),
            avg_pretrend = mean(abs(att))) |> 
  mutate(ratio = overall_att / avg_pretrend,
         modelspec = case_match(modelspec,
                                "queer_nsg_cov" ~ "Cis LGBQ+ Residents of Treated States",
                                "trans_nsg_cov" ~ "Trans Residents of Not-Yet-Treated States")) |> 
  rename("Comparison Group" = modelspec,
         "Overall ATT" = overall_att,
         "Average Pre-Period ATT Magnitude" = avg_pretrend,
         "Ratio" = ratio)

knitr::kable(ratios, digits = 3)
```

## eTable 2. Sample Construction Process

constructed in Word

## eTable 3. Sample Comparisons

## eTable 4. Alternate Specifications

## eTable 5. Sample Characteristics: Trans Sample

## eFigure 1. Panel Structure of Data
