---
title: "Anti-Trans Laws and Mental Health: Tables and Figures"
format: html
editor: visual
execute: 
  echo: false
  warning: false
  message: false
---

# setup

```{r}
#| label: loading in data and packages

library(tidyverse)
library(here)
library(did)
library(patchwork)

load(here("data_clean", "datasets.Rdata"))
load(here("did_results", "coeffs.Rdata"))
load(here("data_clean", "states_sf.rda"))
load(here("data_clean", "all_laws.Rdata"))
load(here("data_clean", "hps_data.Rdata"))

```

```{r}
#| label: data wrangling

# making some relevant variables
df_queer <- map(df_queer, ~drop_na(.x,
                                   anxious, worry, interest, down) |> 
                  mutate(phq4 = anxious + worry + interest + down - 4, #HPS codes these from 1-4; we need them added as though they are 0-3
                          group = case_when(trans_gnc == 1 ~ treatment_period,
                                              trans_gnc == 0 ~ 0), #`did` package convention: comparison group's group is 0
                         coldeg = case_match(educ, 
                                             c(1:4) ~ 0,
                                             c(5:7) ~ 1,
                                             .default = NA_integer_),
                         nhw = case_match(as.numeric(race_rc),
                                          1 ~ 1,
                                          c(2:4) ~ 0,
                                          .default = NA_integer_),
                         age = year(startdate) - birth_year))

#using the same process on the trans datasets, for later
df_trans <- map(df_trans, ~drop_na(.x,
                                   anxious, worry, interest, down) |> 
                  mutate(phq4 = anxious + worry + interest + down - 4,
                         group = case_when(state_type == "treated" ~ treatment_period,
                                           state_type == "notyettreated" ~ 0),
                         coldeg = case_match(educ, 
                                             c(1:4) ~ 0,
                                             c(5:7) ~ 1,
                                             .default = NA_integer_),
                         nhw = case_match(as.numeric(race_rc),
                                          1 ~ 1,
                                          c(2:4) ~ 0,
                                          .default = NA_integer_),
                         age = year(startdate) - birth_year))

```

# Main Results

## Table 1. Sample Characteristics

```{r}
#| label: sample characteristics
# the queer comparison is the main result; making trans vs. trans a supportive supplemental argument

sc <- map(df_queer, ~group_by(.x, trans_gnc, post_period) |> 
  summarize(n = n(),
            mean_phq4 = weighted.mean(phq4, w = pweight),
            sd_phq4 = sd(phq4),
            mean_age = weighted.mean(age, w = pweight),
            sd_age = sd(age),
            prop_nhw = weighted.mean(nhw, w = pweight),
            prop_coldeg = weighted.mean(coldeg, w = pweight),
            prop_fatbirth = weighted.mean(genid_birth - 1, w = pweight)) |> 
  ungroup())
  
sc <- sc |> 
  tibble() |> 
  mutate(analysis = names(sc)) |> 
  unnest(sc) |> 
  pivot_longer(cols = c(n:prop_fatbirth),
               names_to = "stat",
               values_to = "value") |> 
  mutate(trans_gnc = case_match(trans_gnc,
                                0 ~ "queer",
                                1 ~ "trans"),
         post_period = case_match(post_period,
                                  0 ~ "pre",
                                  1 ~ "post")) |> 
  pivot_wider(id_cols = stat,
              names_from = c(trans_gnc, post_period, analysis),
              values_from = value,
              names_sep = "_")

write_csv(sc, file = here("tables_figs", "desctable.csv"))
```

## Figure 1. States Visualization

```{r}
analysis_states <- df_queer |> map(~group_by(.x, state) |> 
  summarize(n = n()) |> 
  pull(state))

mapdata <- states_sf |> 
  mutate(first = st %in% analysis_states$first,
         bathroom = st %in% analysis_states$bathroom,
         medical = st %in% analysis_states$medical,
         sports = st %in% analysis_states$sports)

 hcl.colors(4, palette = "Pastel 1")
 # "#FFC5D0" "#D4D8A7" "#99E2D8" "#D5D0FC"

map_b <- ggplot(data = mapdata) +
  geom_sf(aes(fill = bathroom)) +
  scale_fill_manual(values = c("grey95", "#99E2D8")) +
  theme_void() +
  labs(title = "First Bathroom Law") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

map_m <- ggplot(data = mapdata) +
  geom_sf(aes(fill = medical)) +
  scale_fill_manual(values = c("grey95", "#FFC5D0")) +
  theme_void() +
  labs(title = "First Medical Law") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

map_s <- ggplot(data = mapdata) +
  geom_sf(aes(fill = sports)) +
  scale_fill_manual(values = c("grey95", "#D4D8A7")) +
  theme_void() +
  labs(title = "First Sports Law") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

map_f <- ggplot(data = mapdata) +
  geom_sf(aes(fill = first)) +
  scale_fill_manual(values = c("grey95", "#D5D0FC")) +
  theme_void() +
  labs(title = "First Law Overall") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))


map_b + map_m + map_s + map_f + plot_layout(nrow = 2)


```

## Figure 2. PHQ-4 Trends Visualization

```{r}
#| label: phq-4 visualization

# getting a list of states included in any analysis
states_inany <- tibble(analysis_states) |> 
  unnest_longer(analysis_states) |> 
  unique() |> 
  pull(analysis_states)

# making the picture
phq4_plot <- hps_data |> 
  filter(state %in% states_inany) |> 
  drop_na(queer, trans_gnc) |> 
  mutate(phq4 = anxious + worry + interest + down - 4,
         group = fct(case_when(queer == 0 & trans_gnc == 0 ~ "Cisheterosexual",
                           queer == 1 & trans_gnc == 0 ~ "Cis LGBQ+",
                           queer == 1 & trans_gnc == 1 ~ "Trans/GNC"), levels = c("Trans/GNC", "Cis LGBQ+",                                                           "Cisheterosexual"))) |> 
  group_by(startdate, group) |> 
  summarize(mean_phq = weighted.mean(phq4, w = pweight, na.rm = T)) |> 
  ggplot(aes(x = startdate, y = mean_phq)) +
  geom_line(aes(group = group, color = group)) +
  scale_color_manual(values = hcl.colors(3, palette = "SunsetDark")) +
  scale_x_date(date_labels = "%b %Y") +
  theme_minimal() +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  labs(x = "Date",
       y = "Survey-Weighted Average PHQ-4 Score",
       color = "Group")
  
# saving
ggsave(filename = 
         here("tables_figs", "phq4plot.png"),
       plot = phq4_plot,
       dpi = 300,
       width = 6,
       height = 5,
       units = "in")
```

## Figure 3. DID Results, Overall

```{r}

hcl.colors(4, palette = "Set 2")
# "#00C1B2" "#ACA2EC""#ED90A4" "#ABB150"

overall_did_plot <- coeffs |>
  map(~select(.x, modelspec, overall_att, overall_se) |> 
  unique() |> 
  filter(str_detect(modelspec, "queer"))) |> 
  tibble() |> 
  mutate(law = names(coeffs)) |> 
  unnest(1) |> 
  mutate(modelspec = fct(case_match(modelspec,
                                "queer" ~ "Unadjusted",
                                "queer_nsg" ~ "Removing Small Groups",
                                "queer_cov" ~ "Removing Small Groups,\nAdding Covariates"),
                         levels = c("Removing Small Groups,\nAdding Covariates",
                                    "Removing Small Groups",
                                    "Unadjusted")),
         law = factor(law, levels = c("bathroom", "medical", "sports", "first"), labels = c("First Bathroom Law", "First Medical Care Law", "First Sports Law", "First Law Overall"))) |> 
  ggplot(aes(y = modelspec, color = law)) +
  geom_point(aes(x = overall_att)) +
  geom_segment(aes(x = overall_att + 1.96*overall_se, xend = overall_att - 1.96*overall_se)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap("law") +
  scale_color_manual(values = c("#00C1B2","#ED90A4","#ABB150", "#ACA2EC")) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.background = element_rect(color = "black")) +
  labs(x = "Overall ATT",
       y = "Model Specifications")

ggsave(filename = here("tables_figs",
                       "overall_did_plot.png"),
       plot = overall_did_plot,
       dpi = 300,
       width = 7,
       height = 4.5,
       units = "in")
```

## Figure 4: Dynamic ATTs

```{r}
dynamic_did_plot <- coeffs |>
  map(~select(.x, modelspec, t, att, se)) |> 
  tibble() |> 
  mutate(law = names(coeffs)) |> 
  unnest(1) |> 
  filter(modelspec == "queer_cov",
         t < 8,
         t > -9,
         law %in% c("sports", "first")) |> 
  mutate(law = factor(law, levels = c("sports", "first"), labels = c("First Sports Law", "First Law Overall")),
         pre_period = t < -1) |> 
  ggplot(aes(x = t, color = law,
             alpha = pre_period)) +
  geom_point(aes(y = att)) +
  geom_segment(aes(y = att + 1.96*se, yend = att - 1.96*se)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = -1) +
  facet_wrap("law") +
  scale_color_manual(values = c("#ABB150", "#ACA2EC")) +
  theme_minimal() +
  scale_alpha_manual(values = c(1, .5)) +
  theme(legend.position = "none",
        panel.background = element_rect(color = "black")) +
  labs(x = "Waves Since Treatment",
       y = "Dynamic ATTs")

ggsave(filename = here("tables_figs",
                       "dynamic_did_plot.png"),
       plot = dynamic_did_plot,
       dpi = 300,
       width = 7,
       height = 4.5,
       units = "in")

```

# Supplemental

## eTable 1. Post-Treatment ATTs vs Pre-Treatment ATTs

```{r}
#| label: post vs. pre ratios
coeffs$sports |> 
  filter(t < -1) |> 
  group_by(modelspec) |> 
  summarize(overall_att = mean(overall_att),
            avg_pre = mean(abs(att))) |> 
  mutate(ratio = overall_att / avg_pre)

coeffs$first |> 
  filter(t < -1) |> 
  group_by(modelspec) |> 
  summarize(overall_att = mean(overall_att),
            avg_pre = mean(abs(att))) |> 
  mutate(ratio = overall_att / avg_pre)

```

## eTable 2. Sample Construction Process

constructed in Word

## eTable 3. Sample Comparisons

## eTable 4. Alternate Specifications

## eTable 5. Sample Characteristics: Trans Sample

## eFigure 1. Panel Structure of Data
