---
title: "BRFSS data analysis"
format: html
editor: visual
---

## Background

I started this project by analyzing Household Pulse Survey data, with the intent of doing a difference-in-differences analysis; unfortunately, parallel trends were not met, and the sample of LGBTQ+ individuals was small enough that most of the result was just noise.

So, now I'm taking a different tack: looking at states over periods of years, and comparing mental health of the population in those years.

This will require BRFSS data.

# 1. BRFSS

```{r}
#| label: setup

library(tidyverse)
library(here)
library(rvest)
library(janitor)
library(haven)
library(pdftools)

if (dir.exists(here("codebooks"))) {
  message("Codebooks will be saved in ", here("codebooks"))
} else {
  dir.create(here("codebooks"))
  message("Codebooks will be saved in ", here("codebooks"))
}

if (dir.exists(here("data_raw"))) {
  message("Raw datafiles will be saved in ", here("data_raw"))
} else {
  dir.create(here("data_raw"))
  message("Raw datafiles will be saved in ", here("data_raw"))
}

if (dir.exists(here("data_clean"))) {
  message("Cleaned datafiles will be saved in ", here("data_clean"))
} else {
  dir.create(here("data_clean"))
  message("Cleaned datafiles will be saved in ", here("data_clean"))
}
```

```{r}
#| label: fetching htmls

# the BRFSS pages have a common url pattern, which makes this easier
# i'm using data from 2016 to 2023

urls <- tibble(years = c(2016:2023)) |> 
  mutate(urls = paste0("https://www.cdc.gov/brfss/annual_data/annual_", years, ".html"))

htmls <- urls$urls |> 
  set_names() |> 
  map(~read_html(.x))
```

```{r}
#| label: downloading zipfiles

# the data come in ascii format and SAS xpt format; I find the xpt easier to work with, so...

get_xpt_link <- function(html) {
 xpt_link <-  html |> 
    html_elements("a") |> 
    html_attr("href") |> 
    str_subset("XPT.zip")
 return(xpt_link)
}

zipfiles <- map(htmls, ~get_xpt_link(.x)) |> 
  tibble() |> 
  unnest_longer(1) |> 
  rename(urlend = 1) |> 
  mutate(ziplinks = paste0("https://www.cdc.gov", urlend),
         filename = basename(ziplinks))

map2(.x = zipfiles$ziplinks, .y = zipfiles$filename,
     ~curl::curl_download(url = .x, 
                          destfile = here("data_raw", .y), 
                          quiet = FALSE))
```

```{r}
#| label: unzipping data

# first list the contents, make sure there's not multiple files in there
map(here("data_raw", dir(here("data_raw"))),
    ~unzip(.x, list = TRUE))

# now unzip

map(here("data_raw", dir(here("data_raw"))),
    ~unzip(.x, exdir = here("data_raw")))

# now you can get rid of the zipfiles

here("data_raw", dir(here("data_raw"))) |> 
  str_subset(".zip") |> 
  map(~file.remove(.x))

# check it

dir(here("data_raw"))

# a quirk about the BRFSS data is that the XPT files have a space at the end of their filepath; here's a way to remove it

filename_fix <- tibble(old = dir(here("data_raw"))) |> 
  mutate(new = str_squish(old))


map2(.x = filename_fix$old, .y = filename_fix$new,
     ~file.rename(from = paste0(here("data_raw"), "/", .x), 
                  to = paste0(here("data_raw"), "/", .y)))


# check
dir(here("data_raw"))
```

```{r}
#| label: codebooks

# in order to read in only the data we need, we need the codebooks


get_codebook_link <- function(html) {
  codebook <- html |> 
    html_elements("a") |> 
    html_attr("href") |> 
    str_subset("codebook")
  return(codebook)
}

codebooks <- map(htmls, ~get_codebook_link(.x)) |> 
  tibble() |> 
  unnest_longer(1) |> 
  rename(urlend = 1) |> 
  mutate(codelinks = paste0("https://www.cdc.gov", urlend),
         filenames = basename(codelinks))

map2(.x = codebooks$codelinks,
     .y = codebooks$filenames,
     ~curl::curl_download(.x, destfile = here("codebooks", .y),
                          quiet = FALSE))

#2022 & 2023 are zipfiles

map(here("codebooks", 
         str_subset(dir(here("codebooks")), ".zip")),
    ~unzip(.x, exdir = here("codebooks")))
    
# cleaning up

here("codebooks", dir(here("codebooks"))) |> 
  str_subset(".zip") |> 
  map(~file.remove(.x))

codebook_fix <- tibble(old = dir(here("codebooks"))) |> 
  mutate(beginning = str_extract(old, "([0-9]+)"),
         end = str_extract(old, "\\.[A-Za-z]+"),
         new = paste0("codebook", beginning, end))

map2(.x = paste0(here("codebooks"), "/", codebook_fix$old),
.y = paste0(here("codebooks"), "/", codebook_fix$new),
~file.rename(from = .x, to = .y))
```

```{r}
#| label: reading in the data

datalist <- here("data_raw", dir(here("data_raw"))) |> 
  set_names(nm = basename) |> 
  map(~read_xpt(.x,
         col_select = any_of(c("_STATE",
             "_RACE",
             "_RACE1",
             "_AGE80",
             "SEX",
             "SEX1",
             "SXORIENT",
             "SOMALE",
             "SOFEMALE",
             "SEXVAR",
             "TRNSGNDR",
             "EDUCA",
             "MARITAL",
             "GENHLTH",
             "MENTHLTH",
             "PHYSHLTH",
             "POORHLTH",
             "_LLCPWT"))))
```

```{r}
#| label: data cleaning

datalist <- map(datalist, ~zap_label(.x)) # saves a bunch of space

# 2016 and 2017 need to be cleaned separately from the others, bc they have a different number of variables

datalist1 <- datalist |> 
  keep(str_detect(names(datalist), "(2016)|(2017)"))

datalist2 <- datalist |> 
  discard(str_detect(names(datalist), "(2016)|(2017)"))


# first one first

data1 <- map(datalist1, clean_names) |> 
  map2(.y = names(datalist1),
       ~mutate(.x, source = .y)) |> 
  bind_rows() |>
  mutate(across(c(physhlth, menthlth, poorhlth), #mental/physical health days
                ~case_match(.x,
                           88 ~ 0,
                           c(77,99) ~ NA_integer_,
                           .default = .x)),
         genhlth = case_match(genhlth,
                              c(7,9) ~ NA_integer_,
                              c(1:5) ~ 6 - genhlth,
                              .default = genhlth),# reverse coding so higher is better health
         female = case_match(sex,
                             9 ~ NA_integer_,
                             2 ~ 1,
                             1 ~ 0), # making a binary variable for sex
         marital = case_match(marital,
                              c(1,6) ~ "marpar",
                              c(2:4) ~ "divwidsep",
                              5 ~ "nevmar",
                              9 ~ NA_character_,
                              .default = as.character(marital)),
         # collapsing into 3 categories
         educa = case_match(educa,
                            c(1:3) ~ "lths",
                            4 ~ "hsgrad",
                            5 ~ "somcol",
                            6 ~ "colgrad",
                            9 ~ NA_character_,
                            .default = as.character(educa)),
         # collapsing into 4 categories
         sxorient = case_match(sxorient,
                               1 ~ "st",
                               2 ~ "gay",
                               3 ~ "bi",
                               4 ~ "oth",
                               c(7,9) ~ NA_character_,
                             .default = as.character(sxorient)),
         # recoding to character variable
         tgnc = case_match(trnsgndr,
                           c(1:3) ~ 1,
                           4 ~ 0,
                           c(7,9) ~ NA_integer_,
                           .default = trnsgndr),
         # making a binary variable: trans/gender non-conforming or not
         race = case_match(race,
                           1 ~ "nhw",
                           2 ~ "nhb",
                           c(3:7) ~ "othmult",
                           8 ~ "hisp",
                           9 ~ NA_character_),
         # collapsing to 4 categories
         age = age80,
         state_code = state) |>
  select(state_code, source, sxorient, tgnc, 
         marital, female, race, age, educa,
         genhlth, physhlth, menthlth, poorhlth) |> 
# originally 936,319 observations
  drop_na(sxorient)


## second one 

rename_sex <- function(df) {
  sexvar <- colnames(df) |> str_subset("SEX")
  
  df <- df |> 
    rename(sex = all_of(sexvar))
  
  return(df)
} # need to make sure all of these 

data2 <- map(datalist2, rename_sex) |> 
  map(clean_names) |> 
  map2(.y = names(datalist2),
       ~mutate(.x, source = .y)) |> 
  bind_rows() |>
  mutate(across(c(physhlth, menthlth, poorhlth), #mental/physical health days
                ~case_match(.x,
                           88 ~ 0,
                           c(77,99) ~ NA_integer_,
                           .default = .x)),
         genhlth = case_match(genhlth,
                              c(7,9) ~ NA_integer_,
                              c(1:5) ~ 6 - genhlth,
                              .default = genhlth),# reverse coding so higher is better health
         female = case_match(sex,
                             9 ~ NA_integer_,
                             2 ~ 1,
                             1 ~ 0), # making a binary variable for sex
         marital = case_match(marital,
                              c(1,6) ~ "marpar",
                              c(2:4) ~ "divwidsep",
                              5 ~ "nevmar",
                              9 ~ NA_character_,
                              .default = as.character(marital)),
         # collapsing into 3 categories
         educa = case_match(educa,
                            c(1:3) ~ "lths",
                            4 ~ "hsgrad",
                            5 ~ "somcol",
                            6 ~ "colgrad",
                            9 ~ NA_character_,
                            .default = as.character(educa)),
         # collapsing into 4 categories
         sxorient = case_when(sex == 2 ~ sofemale,
                              sex == 1 ~ somale),
         sxorient = case_match(sxorient,
                               1 ~ "gay",
                               2 ~ "st",
                               3 ~ "bi",
                               4 ~ "oth",
                               c(7,9) ~ NA_character_,
                             .default = as.character(sxorient)),
         # recoding to character variable
         tgnc = case_match(trnsgndr,
                           c(1:3) ~ 1,
                           4 ~ 0,
                           c(7,9) ~ NA_integer_,
                           .default = trnsgndr),
         # making a binary variable: trans/gender non-conforming or not
         race = case_match(race,
                           1 ~ "nhw",
                           2 ~ "nhb",
                           c(3:7) ~ "othmult",
                           8 ~ "hisp",
                           9 ~ NA_character_),
         # collapsing to 4 categories
         age = age80,
         state_code = state) |> 
  select(state_code, source, sxorient, tgnc, 
         marital, female, race, age, educa,
         genhlth, physhlth, menthlth, poorhlth) |> 
  # had 2574810 observations
  drop_na(sxorient)


brfss <- bind_rows(data1, data2) |> 
  mutate(year = as.numeric(str_extract(source, "[0-9]+"))) |> 
  relocate(year, .after = "state_code") |> 
  relocate(source, .after = everything())

```

```{r}
#| label: saving cleaned brfss data

save(brfss, file = here("data_clean", "brfss.Rdata"))

```

once you've checked to make sure that the BRFSS data is saved, you can run the chunk below to clean up. CHECK FIRST!!!!!!! otherwise you'll have to do most of this over again.

```{r}
#| label: halftime cleanup

rm(list = ls())
gc()

if (dir.exists(here("data_raw", "brfss"))) {
  message("BRFSS datafiles will be moved to ", here("data_raw", "brfss"))
} else {
  dir.create(here("data_raw", "brfss"))
  message("BRFSS datafiles will be moved to ", here("data_raw", "brfss"))
}

movefiles <- dir(here("data_raw")) |> str_subset(".XPT")

for (file in movefiles) {
  old_path <- here("data_raw", file)
  new_path <- here("data_raw", "brfss", file)
  
  file.copy(from = old_path,
            to = new_path)
  
  file.remove(here("data_raw", file))
}
```

# 2. Anti-Trans Laws

I'm getting data from the Movement Advancement Project (MAP).

```{r}
#| label: setup for round two

if (dir.exists(here("data_raw", "laws"))) {
  message("Laws PDFs will be saved in ", here("data_raw", "laws"))
} else {
  dir.create(here("data_raw", "laws"))
  message("Laws PDFs will be saved in ", here("data_raw", "laws")) }
```

```{r}
#| label: data acquisition

pdfs <- tibble(url = c("https://www.lgbtmap.org/img/maps/citations-bathroom-facilities-bans.pdf", 
               "https://www.lgbtmap.org/img/maps/citations-sports-participation-bans.pdf", 
               "https://www.lgbtmap.org/img/maps/citations-youth-medical-care-bans.pdf")) |> 
  mutate(filename = basename(url))


map2(.x = pdfs$url, .y = pdfs$filename,
     ~curl::curl_download(url = .x,
                          destfile = here("data_raw", 
                                          "laws", .y),
                          quiet = FALSE))

pdf_parser <- function(file) {
    
  page1 <- NULL
  page2 <- NULL
  text <- pdf_text(here("data_raw", "laws", file))
  
    for (i in 2:length(text)) {
      if (str_detect(text[i], "Order of Law")) {
        page1 <- i
        break
      }
    }
      
    for (i in 2:length(text)) {
      if (str_detect(text[i], "Order of Governor Vetoes")) {
        page2 <- i
        break
      }
    }
     
  return(text[page1:page2])
  
    }
  

text <- dir(here("data_raw", "laws")) |> 
  set_names() |> 
  map(pdf_parser)

```

```{r}
#| label: data cleaning
laws_data <- map(text, ~read_delim(.x, delim = "\n") |> 
  rename(data = 1) |> 
  mutate(data = str_squish(data)) |>
  filter(!str_detect(data, "^20")) |> 
  mutate(drop = if_else(str_detect(data, "Order of Governor Vetoes"), 1, 0),
         drop = cumsum(drop)) |> 
  filter(drop == 0) |> 
  mutate(drop = if_else(str_detect(data, "^1."), 0, NA_integer_)) |> 
  fill(drop, .direction = "down") |> 
  drop_na() |> 
  select(-drop) |> 
  separate(data, into = c("state", "bill", "date"), 
           sep = "[–—]") |> 
  mutate(state = str_remove_all(state, "^([[:digit:]]|[[:punct:]]|[[:space:]])+")) |> 
  separate(date, into = c("date", "notes"), sep = "\\(") |> 
  mutate(across(everything(), ~str_squish(.x)),
         date = mdy(date)))


laws <- tibble(laws_data) |> 
  mutate(pdf_name = names(laws_data)) |> 
  unnest(laws_data)
```

```{r}
#| label: saving laws data
save(laws, file = here("data_clean", "laws.Rdata"))

```

```{r}
#| label: cleanup

rm(list = ls())
gc()

```
